---
title: "Data Cleaning"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# These are Global Options for Your Document
knitr::opts_chunk$set(echo = TRUE)
```

# Wide to Long, Long to Wide

As discussed in the slides, we will use functions from `tidyr` (automatically loaded with the `tidyverse`) to manipulate our data.

## `gather` for wide to narrow

Let's read in our data set.

First we need to call in our libraries:

```{r library}
library(tidyverse)
```

Now we can read in our data using `readr`'s `read_csv` function

```{r read_wide}
gap_wide <- read_csv("data/gapminder_wide.csv")
```

Go ahead and look at the data in the environment pane

## Tidy it up

Now we need to use the `gather` function:

```{r make_narrow}
gap_narrow <- gap_wide %>% 
  gather(key = "country", value = "population", -year)
```

Now we can inspect our work:

```{r inspect_narrow}
head(gap_narrow)
```

As a teaser, this can be used to make some very slick graphics

```{r fact_plot}
gap_narrow %>% 
  filter(country %in% c("India", "China", "Germany", "France")) %>% 
  ggplot(aes(year, population))+
  geom_smooth()+
  facet_wrap(~country)
```

# Now To Spread with `spread`

Sometimes data come in a long format where it needs to be a little wider to fit our tidy paradigm.

Let's experiment with spreading data. Let's read in the `health_long.xlsx` file from our data folder.

This requires the `readxl` package.

```{r readxl}
library(readxl)
```

Now let's read it into memory.

```{r read_long}
health_long <- read_excel("data/health_long.xlsx")
```

We can look at the format of the data

```{r preview_long}
head(health_long)
```

It appears that each subject appears in multiple rows with repeating attributes in the "measurement" column.

Thus we want to spread *measurement* (which will make new column names) and then have the values in these new columns be those values in the *value* column. 

```{r}
health_wide %>% 
  spread(key = measurement, value = value)
```

# Exercise 1 - Your Turn!

Read in the `heights.dta` file. This file contains data on some physical attributes and earnings of subjects in a study.

Hint: you will need to use the `haven` package. Put 

## read in the data

```{r}
library("package")
heights_raw <- read_****("data/")
```

## tidy it

```{r}
heights_raw %>% 
  verb()
```

# Exercise 2