---
title: "Data Cleaning"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# These are Global Options for Your Document
knitr::opts_chunk$set(echo = TRUE)
```

# Wide to Long, Long to Wide

As discussed in the slides, we will use functions from `tidyr` (automatically loaded with the `tidyverse`) to manipulate our data.

## `gather` for wide to narrow

Let's read in our data set.

First we need to call in our libraries:

```{r library}
library(tidyverse)
```

Now we can read in our data using `readr`'s `read_csv` function

```{r read_wide}
gap_wide <- read_csv("data/gapminder_wide.csv")
```

Go ahead and look at the data in the environment pane

## Tidy it up

Now we need to use the `gather` function:

```{r make_narrow}
gap_narrow <- gap_wide %>% 
  gather(key = "country", value = "population", -year)
```

Now we can inspect our work:

```{r inspect_narrow}
head(gap_narrow)
```

As a teaser, this can be used to make some very slick graphics

```{r fact_plot}
gap_narrow %>% 
  filter(country %in% c("India", "China", "Germany", "France")) %>% 
  ggplot(aes(year, population))+
  geom_smooth()+
  facet_wrap(~country)
```

# Now To Spread with `spread`

Sometimes data come in a long format where it needs to be a little wider to fit our tidy paradigm.

Let's experiment with spreading data. Let's read in the `health_long.xlsx` file from our data folder.

This requires the `readxl` package.

```{r readxl}
library(readxl)
```

Now let's read it into memory.

```{r read_long}
health_long <- read_excel("data/health_long.xlsx")
```

We can look at the format of the data

```{r preview_long}
head(health_long)
```

It appears that each subject appears in multiple rows with repeating attributes in the "measurement" column.

Thus we want to spread *measurement* (which will make new column names) and then have the values in these new columns be those values in the *value* column. 

```{r}
health_long %>% 
  spread(key = measurement, value = value)
```

# Exercise 1 - Your Turn!

Read in the `heights.dta` file. This file contains data on some physical attributes and earnings of subjects in a study.

Hint: you will need to use the `haven` package. Put the data into a tidy format. You will note that it is currently in a `long` format.

## read in the data

```{r}
library(haven)
heights_raw <- read_dta("data/heights.dta")
```

## tidy it

Long -> Wide

```{r}
heights_raw %>% 
  spread(key = description, value = value, fill = NA)
```

# Exercise 2

Read in the `gapminder.sav` data set and collapse all the metrics for life expectancy, population and GDP per Population into two columns one for `parameter_name` and the other for `value`.

```{r}
gapminder<-read_sav("data/gapminder.sav")
```

Hint: you will need to use the `-group_1, -group_2, etc` syntax to not collapse the grouping variables that you wish to keep. 

```{r}
gapminder %>% 
  gather(key = parameter_name, value = value, -country, -continent, -year)
```


# Excerise 3

Using `select` to subset

From the `gapminder` data set that you already read into memory, select the "year" and "pop" columns

```{r}
gapminder %>% 
  select(year, pop)
```


# Excerise 4

Use the `unite` function to combine "year" and "country" into one column called country_year with values separated by a "-". Save this into an object called "unite_demo."

```{r}
# In this case I will specify the two columns I want to joinh
# Equivalently I could have used the -lifeExp, -Continent, -pop, -gpdPercap and gotten the same results
unite_demo <- gapminder %>% 
  unite(country_year, sep = "-", c(year, country))

unite_demo
```

# Exercise 5

Now convert this new data set back into the "year" and "country" columns using `separate`.

```{r}
unite_demo %>% 
  separate(col = country_year, into = c("country", "year"), sep = "-")
```


# Exercise 6

Now let's return to the `gap_narrow` dataset and filter the "year" field for values for "1977" only using the `filter` function and the `==` operator

```{r}
gap_narrow %>% 
  filter(year == 1977)
```

# Bonus

Using the `gap_narrow` dataset let's print the row with the maximum population.


# Exercise 7

Let's combine a few operations together using the pipe `%>%`.

- Take the "health_long" data set

- `spread` to a tidy format

- `rename` the "subject" column to "subject_id"

- use `mutate` to create a value for total cholesterol (e.g. total_cholesterol = lhl + hdl)

```{r}
health_long %>% 
  spread(key = measurement, value = value, fill = NA) %>% 
  rename(subject_id = subject) %>% 
  mutate(total_cholesterol = ldl + hdl)
```

Of course if I want to write this data out I can do so:

```{r}
my_new_health_data <- health_long %>% 
  spread(key = measurement, value = value, fill = NA) %>% 
  rename(subject_id = subject) %>% 
  mutate(total_cholesterol = ldl + hdl)

write_csv(my_new_health_data, "outputs/my_new_health_data.csv")
```


# Bonus

I discussed `group_by` for group-wise operations. Taking our gapminder data we could summarise the world population by year.

```{r}
gapminder %>% 
  group_by(year) %>%
  summarise(total_pop = max(pop))
```

Or by year and continent:

```{r}
gapminder %>% 
  group_by(year, continent) %>%
  summarise(total_pop = max(pop))
```

This then lets us do some neat graphing

```{r}
gapminder %>% 
  group_by(year, continent) %>%
  summarise(total_pop = max(pop)) %>% 
  ggplot(aes(year, total_pop, 
             group = continent, color = as_factor(continent)))+
  geom_line()+
  scale_y_log10()+
  labs(
    title = "Population Trends Over Time",
    subtitle = "From The Hans Rosling's Gapminder Dataset",
    y = "Population (log10)",
    x = NULL,
    color = "Continent"
  )+
  theme_minimal()
```

