---
title: "Dates and Times"
author: "Michael DeWitt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

The purpose of this exercise is to explore some of the functionality of the `lubridate` package to help us handle dates and times with easy.

# Packages

LEt's call our required packages. In addition to the `tidyverse` we will need to use the `lubridate` package.

```{r}
library(tidyverse)
library(lubridate)
```


# Dealing With Dates

Dates often arrive in different formats (because few people realise that there are [ISO standards](https://en.wikipedia.org/wiki/ISO_8601) for dates). 

## Some Traditional Formats

1. YYYY-MM-DD

1. MM-DD-YYYY

1. MM/DD/YYYY


Let's make some examples dates:
```{r}
date_1 <- "2010-10-20"

date_2 <- "11-05-1998"

date_3 <- "11/05/1998"

date_4 <- "1/05/2013"
```


### Parsing Each Date

YYYY-MM-DD is by far the easiest date representation (because it is the ISO standard format). Let's use the `ymd` function from lubridate to convert this object into a date.

```{r}
date_1 <- ymd(date_1)
```

Now let's check if it worked!

```{r}
class(date_1)
```

It did! Yay!

Now Let's Try a few others

```{r}

```

And Check

```{r}
class(date_2)
```

```{r}

```

And Check

```{r}
class(date_2)
```


# Parsing Pieces of Dates

Let's practice parsing our dates. 

The functions we want to practice are:

`day`
`wday`
`month`
`week`
`year`

Let's Parse

First we need to convert our "date" column into a date type column

```{r}
lakers

```

Now that we have done that, let's Retrive the components

```{r}

```

Then lets add the months, days and years

```{r}

```

Now let's use our data summarisation tools and see the average points by day of the week by month!

```{r}

```


# Now on to Time

Let's add a time column to our data set


# Now let's calculate some intervals...


