---
title: "Demonstraining Factors"
author: "Michael DeWitt"
date: "11/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
```

# Introducing Forcats

`forcats` is a package that is designed to mke working with factors (if not fun) easy

Factors provide us with a way for handling categorical variables

Additionally, factors make graphing data easier (especially in regards to ordering values)

# Quick Aside

There are some nice templates available for graphing from the `papaja`package. These are APA templates. Go ahead and install them if you are interested. 

```{r eval=FALSE}
devtools::install_github("crsh/papaja")
```

# Brining in the Data

```{r}
gss_cat
```


# Doing Some Modeling

## The Formula Syntax

R uses a specific formula syntax for building models.

`Depedent Variable ~ Indepedent Variable`

This syntax is generally consistent across all modeling functions (exceptions include some tree based methods)

We can add main effects with a `+` and internation effects with `:`. We can expand indepdent variables to include main and interaction effects with `*` (e.g. ($predictor1*predictor2 = predictor1 + predictor2 + predictor1*predictor2$))

## A Linear Model
$$tvhours = \beta_1*tvhours + \beta_2*partyid + \alpha + \epsilon$$

```{r}
fit_1 <- lm(age ~ tvhours + partyid, data = gss_cat)
```

## Now Let's Examine The Output

`summary` will return some of the summary statistics for our model fit

```{r}
summary(fit_1)
```

Yuck! 

## Let's use some of `forcats` functions

Reduce partyid to 3 categories?

```{r}

```

## Simplified Linear Fit

```{r}

```


## Now Some Generalised Linear Modeling

Here we can specify the generalised linear model family we want to use in the `glm` function with the `family` argument. If we don't specify a family it will use the default of `gaussian` or a basic linear regression.

```{r}

```


```{r}

```


## Now some hierarchical modeling

```{r}
library(lme4)
```


```{r}

```


# Dealing with external data

Now an example where we get an `sav` file from NORC for the actual GSS. In this example you will see that it comes in with numbers rather than the labels. This can be problematic.

```{r}
library(haven)
# Purpose: import data from 2016 GSS and format for R

url <- "http://users.wfu.edu/dewittme/data/sample/GSS2016.sav"

# Read Raw Data
raw_data <-read_spss(url)
```

Uh oh...

```{r}
head(raw_data)
```

But what does this <s3: labelled> mean???

```{r}
attributes(raw_data$MAR1)
```

```{r}
clean_data_1 <- labelled::to_factor(raw_data)
clean_data_1 %>% head()
```

```{r eval=FALSE}
library(codebook)
codebook::codebook(raw_data)
```

# Graphing

```{r}
library(papaja)
gss_cat %>% 
  ggplot(aes(rincome))+
  geom_bar()+
  labs(
    title = "Missing Income Data Is High",
    subtitle = "How does this effect our conclusions?",
    caption = "Data: General Social Survey",
    x= "income",
    y = "count"
  )+
  theme_apa()+
  coord_flip()
```

```{r}
gss_cat %>% 
  mutate(relig = fct_lump(relig, n = 5)) %>% # Lump into five categories
  mutate(relig = fct_inorder(relig)) %>%  # Reorder from most to least
  ggplot(aes(relig, tvhours))+
  geom_jitter()+
  labs(
    title = "Missing Income Data Is High",
    subtitle = "How does this effect our conclusions?",
    caption = "Data: General Social Survey",
    x= "income",
    y = "count"
  )
```

