---
title: "ggplot2"
author: "Michael DeWitt"
date: "11/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to `ggplot2`

This is a follow-along guide for the slides in the presentation.

# ggplot2

`ggplot2` is an implementation of the [Grammar of Graphics](https://www.amazon.com/Grammar-Graphics-Statistics-Computing/dp/0387245448). It is not a guide on what makes a _good_ graphic, but rather a philosophy of building statistical graphics. The grammar proposes components that are common to all statistical graphics:

* **data** - what data are you trying to plot

* **mappings** - what aesthic mapping are you trying to plot

* **layers**
  * geometries - bars, lines, points, etc
  * statistics - statistical summaries of the data (e.g. counts)
  
* **scales** - map color, size, shape

* **coordinate systems** - describe how the data are mapped

* **facets** - breaking the data into subsets of small multiples

* **themes** - the details of display like font size, color pallets, etc

# Now for implementing the grammar with `ggplot2`

Let's call the `tidyverse` which includes `ggplot2`.


```{r}
library(tidyverse)
```

We'll start with using the `diamonds` dataset that is built into `ggplot2`[^1].

[^1]: Find more about this dataset using help("diamonds").

```{r}
head(diamonds)
```

This dataset contains information for a variety of diamonds which includes the 5 C's and additional geometries.[^2]

[^2]: See more [about the five cees](https://www.portfoliooffinediamonds.com/Blog/March-2015/The-5-Cs-of-a-diamond)

# Our First Example

## Data
Start with the data:
```{r}
ggplot()
```

## Aesthetics

## Layers

## Scales

## Facets

## Theme Elements

## Saving


